---
title: "Site Clinical Characteristics"
author: "Alex Baras"
date: "November 26, 2016"
output: html_document
---


```{r, echo=FALSE}

# libraries
library(synapseClient)
library(knitr)
library(lattice)

# login to SAGE
#synapseLogin(username="",password="") # you could do this with prompt or coded, but will need to enter something here

# get clinical data file from SAGE
genieClinData = read.delim(synGet("syn7392892")@filePath)

```

# Race

### Counts - note that some entries are still being submitted as "" and not mapped to Unknown

```{r, echo=FALSE}

# Center vs Race TABLE
t = as.data.frame.matrix(table(genieClinData$CENTER,genieClinData$PRIMARY_RACE))
t = data.frame(n=rowSums(t),t)
colnames(t)[2] = "submitted as empty"
kable(t)

```

### Percent unknown ("" is mapped to Unknown for this analysis)

```{r, echo=FALSE}

t[,"Unknown"] = t[,"Unknown"]+t[,2]
t = t[,-2]
kable(data.frame("Percent.Unknown"=(t[,"Unknown"]/rowSums(t[,colnames(t)!="n"]))*100),digits=2)

```

### breakdown for sites with greater than 100 patients with none "Unknown" data

```{r, echo=FALSE}

t$n = t$n-t$Unknown
t = t[,!(colnames(t)=="Unknown")]
k = which(t$n>100)
t = t[,c(1,6,3,2,4,5)]
t[k,colnames(t)!="n"] = t[k,colnames(t)!="n"]/rowSums(t[k,colnames(t)!="n"])
kable(t[k,],digits=3)

```

### Race barplots

```{r, echo=FALSE}

# Center vs Race BARCHARTS
t$Center = rownames(t)
l = colnames(t)[2:6]
t = reshape(t[k,],direction="long",idvar=("Center"),varying=l,v.names="Percentage",times=l,drop="n")
colnames(t)[colnames(t)=="time"] = "Race"
t$Race = factor(t$Race,l)
barchart( t$Race ~ t$Percentage | t$Center,xlim=c(0,1))

```

# Ethnicity

### raw counts - note that some entries are still being submitted as "" and not mapped to Unknown

```{r, echo=FALSE}

# Center vs Ethnicity TABLE
t = as.data.frame.matrix(table(genieClinData$CENTER,genieClinData$ETHNICITY))
t = data.frame(n=rowSums(t),t)
colnames(t)[2] = "submitted as empty"
kable(t)

```

### Percent unknown ("" is mapped to Unknown for this analysis)

```{r, echo=FALSE}

t[,"Unknown"] = t[,"Unknown"]+t[,2]
t = t[,-2]
kable(data.frame("Percent.Unknown"=(t[,"Unknown"]/rowSums(t[,colnames(t)!="n"]))*100),digits=2)

```

### breakdown for sites with greater than 100 patients with none "Unknown" data

```{r, echo=FALSE}

t$n = t$n-t$Unknown
t = t[,!(colnames(t)=="Unknown")]
k = which(t$n>100)
t[k,colnames(t)!="n"] = t[k,colnames(t)!="n"]/rowSums(t[k,colnames(t)!="n"])
kable(t[k,],digits=3)

```

### Ethnicity barplots

```{r, echo=FALSE}

# Center vs Ethnicity BARCHARTS
t$Center = rownames(t)
l = colnames(t)[2:3]
t = reshape(t[k,],direction="long",idvar=("Center"),varying=l,v.names="Percentage",times=l,drop="n")
colnames(t)[colnames(t)=="time"] = "Ethnicity"
t$Ethnicity = factor(t$Ethnicity,l)
barchart( t$Ethnicity ~ t$Percentage | t$Center,xlim=c(0,1))

```

# Sex

### raw counts

```{r, echo=FALSE}

# Center vs Sex TABLE
t = as.data.frame.matrix(table(genieClinData$CENTER,genieClinData$SEX))
t = data.frame(n=rowSums(t),t)
kable(t)

```

### sex breakdown across sites

```{r, echo=FALSE}

t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t,digits=3)

```

### sex barplots

```{r, echo=FALSE}

# Center vs Sex BARCHARTS
t$Center = rownames(t)
l = colnames(t)[2:3]
t = reshape(t,direction="long",idvar=("Center"),varying=l,v.names="Percentage",times=l,drop="n")
colnames(t)[colnames(t)=="time"] = "Sex"
t$Sex = factor(t$Sex,l)
barchart( t$Sex ~ t$Percentage | t$Center,xlim=c(0,1))

```

# Age

### histograms for age across sites

```{r, echo=FALSE}

# Center vs Sex histograms

genieClinData$age_days = as.numeric(levels(genieClinData$AGE_AT_SEQ_REPORT))[genieClinData$AGE_AT_SEQ_REPORT]
k = grep("^>",genieClinData$AGE_AT_SEQ_REPORT) # these are the >=90 group that get mapped to 90
genieClinData$age_days[k] = 90*365
k = grep("^<",genieClinData$AGE_AT_SEQ_REPORT) # these are the <=18 group that get mapped to 18
genieClinData$age_days[k] = 18*365

```

### site breakdown for age at seq reporting not submitted ("" is reason for NA introduced)

```{r, echo=FALSE}

k = is.na(genieClinData$age_days)
t = as.data.frame.matrix(table(genieClinData$CENTER,k))
kable(data.frame("Percent.Unknown"=(t[,"TRUE"]/rowSums(t))*100),digits=2)

```

### site histograms for age at seq reporting histograms

anything starting with ">" is mapped to to 90-108 (18 year bin)

(it looks like SAGE has already age redacted the >=90y bin across site)


```{r, echo=FALSE}

k = !is.na(genieClinData$age_days)
histogram(~(genieClinData$age_days[k]/365)|genieClinData$CENTER[k],type="density",breaks=c(0,seq(1,89,2),90,108),right=FALSE)

```

### site breakdown for age at seq reported at as days <18y (numeric value)

(there are none for >90, since it looks like SAGE is arleady taking care of those)


```{r, echo=FALSE}

k = genieClinData$age_days
k[!is.na(genieClinData$age_days)] = "(18y,90y)"
k[is.na(genieClinData$age_days)] = NA
k[genieClinData$age_days<(18*365)] = "<18y"
k[genieClinData$age_days>(90*365)] = ">90y"
t = as.data.frame.matrix(table(genieClinData$CENTER,k))
t$n = rowSums(t)
t = t[,-2]
kable(t)

```

### site breakdown for age at seq reporting histograms

anything starting with ">" is mapped to to 90-108 (18 year bin)

anything starting with "<" is mapped to to 0-18 (18 year bin)

anything with numerical days < 18y is mapped to to 0-18 (18 year bin)

Note: histograms as being presented as density now that bins are not equal size

```{r, echo=FALSE}

genieClinData$age_days[genieClinData$age_days<(18*365)] = 18*365
k = !is.na(genieClinData$age_days)
histogram(~(genieClinData$age_days[k]/365)|genieClinData$CENTER[k],type="density",breaks=c(0,seq(19,89,2),90,108),right=FALSE)

```
