---
title: "Clinical Data Site Characteristics"
author: "Alex Baras"
date: "`r Sys.Date()`"
output: html_document
---


```{r, echo=FALSE}

# libraries
library(synapseClient)
library(knitr)
library(ggplot2)


# login to SAGE
synapseLogin(username=,password=) # you could do this with prompt or coded, but will need to enter something here

# get clinical data file from SAGE
genieClinData = read.delim(synGet("syn7392892")@filePath)

```

# Race

### Counts

```{r, echo=FALSE}

# Center vs Race TABLE
t = as.data.frame.matrix(table(genieClinData$CENTER,genieClinData$PRIMARY_RACE))
t = data.frame(n=rowSums(t),t)
kable(t)

```

### Percent unknown

```{r, echo=FALSE}

kable(data.frame("Percent.Unknown"=(t[,"Unknown"]/rowSums(t[,colnames(t)!="n"]))*100),digits=2)

```

### Breakdown for sites with greater than 100 patients with none "Unknown" data

```{r, echo=FALSE}

# reduce table to none "Unknown" data
t$n = t$n-t$Unknown
t = t[,!(colnames(t)=="Unknown")]
# find sites with greater than 100 nono "Unknown" patients wrt race
k = which(t$n>100)
t = t[,c("n","White","Black","Asian","Native.American","Other")]
t[k,colnames(t)!="n"] = t[k,colnames(t)!="n"]/rowSums(t[k,colnames(t)!="n"])
kable(t[k,],digits=3)

```

### Race barplots

```{r, echo=FALSE}

# Center vs Race BARCHARTS
t$Center = rownames(t)
l = c("White","Black","Asian","Native.American","Other")
t = reshape(t[k,],direction="long",idvar=("Center"),varying=l,v.names="Percentage",times=l,drop="n")
colnames(t)[colnames(t)=="time"] = "Race"
t$Race = factor(t$Race,l)
ggplot(t,aes(Race,Percentage,group=Center))+geom_bar(stat="identity")+facet_wrap(~Center,ncol=4)+ylim(0,1)+coord_flip()

```

# Ethnicity

### Counts

```{r, echo=FALSE}

# Center vs Ethnicity TABLE
t = as.data.frame.matrix(table(genieClinData$CENTER,genieClinData$ETHNICITY))
t = data.frame(n=rowSums(t),t)
kable(t)

```

### Percent unknown

```{r, echo=FALSE}

kable(data.frame("Percent.Unknown"=(t[,"Unknown"]/rowSums(t[,colnames(t)!="n"]))*100),digits=2)

```

### breakdown for sites with greater than 100 patients with none "Unknown" data

```{r, echo=FALSE}

t$n = t$n-t$Unknown
t = t[,!(colnames(t)=="Unknown")]
k = which(t$n>100)
t[k,colnames(t)!="n"] = t[k,colnames(t)!="n"]/rowSums(t[k,colnames(t)!="n"])
kable(t[k,],digits=3)

```

### Ethnicity barplots

```{r, echo=FALSE}

# Center vs Ethnicity BARCHARTS
t$Center = rownames(t)
l = c("Non.Spanish.non.Hispanic","Spanish.Hispanic")
t = reshape(t[k,],direction="long",idvar=("Center"),varying=l,v.names="Percentage",times=l,drop="n")
colnames(t)[colnames(t)=="time"] = "Ethnicity"
t$Ethnicity = factor(t$Ethnicity,l)
ggplot(t,aes(Ethnicity,Percentage,group=Center))+geom_bar(stat="identity")+facet_wrap(~Center,ncol=4)+ylim(0,1)+coord_flip()

```

# Sex

### raw counts

```{r, echo=FALSE}

# Center vs Sex TABLE
t = as.data.frame.matrix(table(genieClinData$CENTER,genieClinData$SEX))
t = data.frame(n=rowSums(t),t)
kable(t)

```

### sex breakdown across sites

```{r, echo=FALSE}

t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t,digits=3)

```

### sex barplots

```{r, echo=FALSE}

# Center vs Sex BARCHARTS
t$Center = rownames(t)
l = c("Female","Male")
t = reshape(t,direction="long",idvar=("Center"),varying=l,v.names="Percentage",times=l,drop="n")
colnames(t)[colnames(t)=="time"] = "Sex"
t$Sex = factor(t$Sex,l)
ggplot(t,aes(Sex,Percentage,group=Center))+geom_bar(stat="identity")+facet_wrap(~Center,ncol=4)+ylim(0,1)+coord_flip()


```

# Age

### histograms for age across sites

```{r, echo=FALSE}

# Center vs Sex histograms

genieClinData$age_days = as.numeric(levels(genieClinData$AGE_AT_SEQ_REPORT))[genieClinData$AGE_AT_SEQ_REPORT]
k = grep("^>",genieClinData$AGE_AT_SEQ_REPORT) # these are the >=90 group that get mapped to 90
genieClinData$age_days[k] = 32872
k = grep("^<",genieClinData$AGE_AT_SEQ_REPORT) # these are the <=18 group that get mapped to 18
genieClinData$age_days[k] = 6573

```

### site breakdown for age at seq reporting not submitted ("" is reason for NA introduced)

```{r, echo=FALSE}

k = is.na(genieClinData$age_days)
t = as.data.frame.matrix(table(genieClinData$CENTER,k))
kable(data.frame("Percent.Unknown"=(t[,"TRUE"]/rowSums(t))*100),digits=2)

```

### site breakdown for age at seq reporting histograms

anything starting with ">" is mapped to 32872 days and represented by an 18 year bin on the histogram

anything starting with "<" is mapped to to 6573 and represented by an 18 year bin on the histogram

SAGE is redacting anything submitted with age <6573 or >32872

Note: histograms as being presented as density now that bins are not equal size


```{r, echo=FALSE}

k = !is.na(genieClinData$age_days)
ggplot(genieClinData[k,],aes(x=(age_days/365),group=CENTER))+geom_histogram(aes(y = (..count..)/width/tapply(..count..,..group..,sum)[..group..]),breaks=c(0,seq(19,89,2),90,108),closed="left")+facet_wrap(~CENTER,ncol=4)

```

# Oncotree breakdowns

```{r, echo=FALSE}

t = as.data.frame.matrix(table(genieClinData$CANCER_TYPE,genieClinData$CENTER))
kable(t)

```
