---
title: "Genomic Data Site Characteristics"
author: "Alex Baras"
date: "`r Sys.Date()`"
output: html_document
---


```{r, echo=FALSE}

# libraries
library(synapseClient)
library(knitr)
library(ggplot2)

# login to SAGE
synapseLogin(username=,password=) # you could do this with prompt or coded, but will need to enter something here

# read aggregated MAF file
genieMutData = read.delim(synGet("syn5571527")@filePath,header=TRUE,sep="\t",quote="",comment="",skip = 1)

# records with count data that preclude a VAF estimate - set VAF to 100% (1/1, alt/depth)
noVAF.idx = which((genieMutData$t_depth==0)|is.na(genieMutData$t_depth))
genieMutData$t_alt_count_num = as.numeric(levels(genieMutData$t_alt_count))[genieMutData$t_alt_count]
genieMutData$t_alt_count_num[noVAF.idx] = 1
genieMutData$t_depth[noVAF.idx] = 1

# germline filter index
nongermline.idx = genieMutData$FILTER!="common_variant"

```

### SUB vs INS vs DEL (pre-germline filter)

```{r, echo=FALSE}

# SUB vs INS/DEL TABLE
t = as.data.frame.matrix(table(genieMutData$Center,genieMutData$Variant_Type))
t = as.data.frame(cbind(rowSums(t[,c("SNP","DNP","TNP","ONP")]),t[,"INS"],t[,"DEL"]))
colnames(t) = c("SUB","INS","DEL")
t = data.frame(n=rowSums(t),t)
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t$SUB),],digits=3)

# SUB vs INS/DEL BARCHARTS
t$Center = rownames(t)
t = reshape(t,direction="long",idvar=("Center"),varying=c("SUB","INS","DEL"),v.names="Percentage",times=c("SUB","INS","DEL"))
colnames(t)[colnames(t)=="time"] = "Type"
t$Type = factor(t$Type,c("SUB","INS","DEL"))
ggplot(t,aes(Type,Percentage,group=Center))+geom_bar(stat="identity")+facet_wrap(~Center,ncol=4)+ylim(0,1)

```

### SUB vs INS vs DEL (post-germline filter)

```{r, echo=FALSE}

# SUB vs INS/DEL TABLE
k = which(nongermline.idx)
t = as.data.frame.matrix(table(genieMutData$Center[k],genieMutData$Variant_Type[k]))
t = as.data.frame(cbind(rowSums(t[,c("SNP","DNP","TNP","ONP")]),t[,"INS"],t[,"DEL"]))
colnames(t) = c("SUB","INS","DEL")
t = data.frame(n=rowSums(t),t)
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t$SUB),],digits=3)

# SUB vs INS/DEL BARCHARTS
t$Center = rownames(t)
t = reshape(t,direction="long",idvar=("Center"),varying=c("SUB","INS","DEL"),v.names="Percentage",times=c("SUB","INS","DEL"))
colnames(t)[colnames(t)=="time"] = "Type"
t$Type = factor(t$Type,c("SUB","INS","DEL"))
ggplot(t,aes(Type,Percentage,group=Center))+geom_bar(stat="identity")+facet_wrap(~Center,ncol=4)+ylim(0,1)

```

### SUB Length (pre-germline filter)

```{r, echo=FALSE}

# SUB TABLES
k = which((genieMutData$Reference_Allele!="-") & (genieMutData$Tumor_Seq_Allele2!="-") & (nchar(as.character(genieMutData$Reference_Allele))==nchar(as.character(genieMutData$Tumor_Seq_Allele2))))
t = as.data.frame.matrix(table(genieMutData$Center[k],nchar(as.character(genieMutData$Reference_Allele[k]))))
t = data.frame(n=rowSums(t),t)
colnames(t) = sub("X","",colnames(t))
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t[,2]),],digits=3)

# SUB HISTOGRAMS
ggplot(genieMutData[k,],aes(nchar(as.character(Tumor_Seq_Allele2)),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),stat="count")+facet_wrap(~Center,ncol=4)+xlim(0,8)

```

### SUB Length (post-germline filter)

```{r, echo=FALSE}

# SUB TABLES
k = which((genieMutData$Reference_Allele!="-")&(genieMutData$Tumor_Seq_Allele2!="-")&(nchar(as.character(genieMutData$Reference_Allele))==nchar(as.character(genieMutData$Tumor_Seq_Allele2)))&nongermline.idx)
t = as.data.frame.matrix(table(genieMutData$Center[k],nchar(as.character(genieMutData$Reference_Allele[k]))))
t = data.frame(n=rowSums(t),t)
colnames(t) = sub("X","",colnames(t))
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t[,2]),],digits=3)

# SUB HISTOGRAMS
ggplot(genieMutData[k,],aes(nchar(as.character(Tumor_Seq_Allele2)),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),stat="count")+facet_wrap(~Center,ncol=4)+xlim(0,8)

```

### INS Length (pre-germline filter)

```{r, echo=FALSE}

# INS TABLE
k = which((genieMutData$Reference_Allele=="-")|(nchar(as.character(genieMutData$Reference_Allele))<nchar(as.character(genieMutData$Tumor_Seq_Allele2))))
t = as.data.frame.matrix(table(genieMutData$Center[k],nchar(as.character(genieMutData$Tumor_Seq_Allele2[k]))))
t = data.frame(n=rowSums(t),t)
colnames(t) = sub("X","",colnames(t))
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t[,2]),],digits=3)

# INS HISTOGRAMS
ggplot(genieMutData[k,],aes(nchar(as.character(Tumor_Seq_Allele2)),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),stat="count")+facet_wrap(~Center,ncol=4)+xlim(0,20)

```

### INS Length (post-germline filter)

```{r, echo=FALSE}

# INS TABLE
k = which((genieMutData$Reference_Allele=="-")|(nchar(as.character(genieMutData$Reference_Allele))<nchar(as.character(genieMutData$Tumor_Seq_Allele2)))&nongermline.idx)
t = as.data.frame.matrix(table(genieMutData$Center[k],nchar(as.character(genieMutData$Tumor_Seq_Allele2[k]))))
t = data.frame(n=rowSums(t),t)
colnames(t) = sub("X","",colnames(t))
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t[,2]),],digits=3)

# INS HISTOGRAMS
ggplot(genieMutData[k,],aes(nchar(as.character(Tumor_Seq_Allele2)),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),stat="count")+facet_wrap(~Center,ncol=4)+xlim(0,20)

```

### DEL Length (pre-germline filter)

```{r, echo=FALSE}

# DEL TABLE
k = which((genieMutData$Tumor_Seq_Allele2=="-")|(nchar(as.character(genieMutData$Reference_Allele))>nchar(as.character(genieMutData$Tumor_Seq_Allele2))))
t = as.data.frame.matrix(table(genieMutData$Center[k],nchar(as.character(genieMutData$Reference_Allele[k]))))
t = data.frame(n=rowSums(t),t)
colnames(t) = sub("X","",colnames(t))
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t[,2]),],digits=3)

# DEL HISTOGRAMS

ggplot(genieMutData[k,],aes(nchar(as.character(Reference_Allele)),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),stat="count")+facet_wrap(~Center,ncol=4)+xlim(0,20)

```


### DEL Length (post-germline filter)

```{r, echo=FALSE}

# DEL TABLE
k = which((genieMutData$Tumor_Seq_Allele2=="-")|(nchar(as.character(genieMutData$Reference_Allele))>nchar(as.character(genieMutData$Tumor_Seq_Allele2)))&nongermline.idx)
t = as.data.frame.matrix(table(genieMutData$Center[k],nchar(as.character(genieMutData$Reference_Allele[k]))))
t = data.frame(n=rowSums(t),t)
colnames(t) = sub("X","",colnames(t))
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t[order(t[,2]),],digits=3)

# DEL HISTOGRAMS

ggplot(genieMutData[k,],aes(nchar(as.character(Reference_Allele)),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),stat="count")+facet_wrap(~Center,ncol=4)+xlim(0,20)

```

### SNV type (pre-germline filter)

```{r, echo=FALSE}

# SNV type breakdown TABLE
k = which(genieMutData$Variant_Type=="SNP")
t = as.data.frame.matrix(table(genieMutData$Center[k],paste(genieMutData$Reference_Allele[k],".",genieMutData$Tumor_Seq_Allele2[k])))
t = data.frame(n=rowSums(t),t)
colnames(t)[2:13] = sub("\\.\\.\\."," -> ",colnames(t)[2:13])
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t,digits=3)

# SNV type breakdown BARCHARTS
t$Center = rownames(t)
t = reshape(t,direction="long",idvar=("Center"),varying=colnames(t)[2:13],v.names="Percentage",times=colnames(t)[2:13],drop="n")
colnames(t)[colnames(t)=="time"] = "Type"
ggplot(t,aes(Type,Percentage,group=Center))+geom_bar(stat="identity")+coord_flip()+facet_wrap(~Center,ncol=4)

```

### SNV type (post-germline filter)

```{r, echo=FALSE}

# SNV type breakdown TABLE
k = which((genieMutData$Variant_Type=="SNP")&nongermline.idx)
t = as.data.frame.matrix(table(genieMutData$Center[k],paste(genieMutData$Reference_Allele[k],".",genieMutData$Tumor_Seq_Allele2[k])))
t = data.frame(n=rowSums(t),t)
colnames(t)[2:13] = sub("\\.\\.\\."," -> ",colnames(t)[2:13])
t[,colnames(t)!="n"] = t[,colnames(t)!="n"]/rowSums(t[,colnames(t)!="n"])
kable(t,digits=3)

# SNV type breakdown BARCHARTS
t$Center = rownames(t)
t = reshape(t,direction="long",idvar=("Center"),varying=colnames(t)[2:13],v.names="Percentage",times=colnames(t)[2:13],drop="n")
colnames(t)[colnames(t)=="time"] = "Type"
ggplot(t,aes(Type,Percentage,group=Center))+geom_bar(stat="identity")+coord_flip()+facet_wrap(~Center,ncol=4)

```

### VAF distribution for SUBs (pre-germline filter)

```{r, echo=FALSE}

k = which(!((genieMutData$Variant_Type=="INS")|(genieMutData$Variant_Type=="DEL")))
ggplot(genieMutData[k,],aes(x=(t_alt_count_num/t_depth),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),breaks=seq(0,1,0.02),closed="left")+facet_wrap(~Center,ncol=4)

```

### VAF distribution for SUBs (post-germline filter)

```{r, echo=FALSE}

k = which(nongermline.idx&!((genieMutData$Variant_Type=="INS")|(genieMutData$Variant_Type=="DEL")))
ggplot(genieMutData[k,],aes(x=(t_alt_count_num/t_depth),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),breaks=seq(0,1,0.02),closed="left")+facet_wrap(~Center,ncol=4)

```

### VAF distribution for INS/DEL (pre-germline filter)

```{r, echo=FALSE}

k = which(((genieMutData$Variant_Type=="INS")|(genieMutData$Variant_Type=="DEL")))
ggplot(genieMutData[k,],aes(x=(t_alt_count_num/t_depth),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),breaks=seq(0,1,0.02),closed="left")+facet_wrap(~Center,ncol=4)

```

### VAF distribution for INS/DEL (pre-germline filter)

```{r, echo=FALSE}

k = which(nongermline.idx&((genieMutData$Variant_Type=="INS")|(genieMutData$Variant_Type=="DEL")))
ggplot(genieMutData[k,],aes(x=(t_alt_count_num/t_depth),group=Center))+geom_histogram(aes(y = (..count..)/tapply(..count..,..group..,sum)[..group..]),breaks=seq(0,1,0.02),closed="left")+facet_wrap(~Center,ncol=4)

```
